Network Visualizations in \MakeLowercase{\texttt{ggplot2}}

### 2015-10-12 inst/paper.Rnw
blood type example cleaned up

\hh{There are two main approaches to making use of the ggplot2 framework: (i) implement network visualizations using ggplot2, i.e. providing a wrapper for the user to visualize a network with ggplot2 elements, and (ii) implement networks as an internal layer of ggplot2. We will discuss both of these approaches in this paper. }

@@

\section{Two approaches of network visualizations}
 The three necessary elements of any network visualization are the vertices, the edges, and the layout.
%But once those three items are visualized, we usually find our visualization to be lacking. \hh{pull it over to a positive statement}  We may want to color the vertices (or edges) by some sort of grouping variable, or we may want to make vertices of degree ten twice the size of vertices of degree five.

\subsection{ggnet}
\hh{XXX needs massive changes :)}
\par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these three necessary graph attributes.
 The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.  Our work builds off of the \texttt{ggnet} function and presents an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.
 \hh{XXXX}

@@

@@

<<blood_ex_ggnet, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
  geom_net(colour = "darkred", layout = "circle", label = TRUE, size = 15,
           directed = TRUE, vjust = 0.5, labelcolour = "grey80", curvature = 0,
           arrowsize = 1.5, linewidth = 0.75, arrowgap = 0.05) +
  theme_net()
@

@@

@@

\begin{figure}
\begin{subfigure}[t]{.49\textwidth}
\caption{geom\_net}
\includegraphics[width=\textwidth]{figure/blood_ex_ggnet-1.pdf}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}
\caption{ggnet2}
\includegraphics[width=\textwidth]{figure/blood_ex_ggnet2-1.pdf}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}
\caption{ggnetwork}
\includegraphics[width=\textwidth]{figure/blood_ex_ggnetwork-1.pdf}
\end{subfigure}
\caption{\label{fig.cap:blood_ex} Network of blood donation possibilities in humans by ABO and RhD blood types.}
\end{figure}
\afterpage{\clearpage}

>>>>

\hh{There are two main approaches to making use of the ggplot2 framework: (i) implement network visualizations using ggplot2, i.e. providing a wrapper for the user to visualize a network with ggplot2 elements, and (ii) implement networks as an internal layer of ggplot2.  For the second approach, two main implementations exist. We will discuss all of these  in this paper. }

@@

\section{Three approaches of network visualizations}
 The three necessary elements of any network visualization are the vertices, the edges, and the layout.
%But once those three items are visualized, we usually find our visualization to be lacking. \hh{pull it over to a positive statement}  We may want to color the vertices (or edges) by some sort of grouping variable, or we may want to make vertices of degree ten twice the size of vertices of degree five.

\subsection{ggnet}
\hh{XXX needs massive changes. Francois, could you take care of that?}
\par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these three necessary graph attributes.
 The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.  Our work builds off of the \texttt{ggnet} function and presents an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.
 \hh{XXXX}

@@

\subsection{geom\_node and geom\_edge}

\hh{Fraancois, could you include a discussion of the background of the ggnetwork pacakge here?}

@@

@@

<<blood_ex_ggnet, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
  geom_net(colour = "darkred", layout = "circle", label = TRUE, size = 15,
           directed = TRUE, vjust = 0.5, labelcolour = "grey80",
           arrowsize = 1.5, linewidth = 0.75, arrowgap = 0.05) +
  theme_net()
@

@@

\begin{figure}
\begin{subfigure}[t]{.32\textwidth}
\caption{ggnet2}
\includegraphics[width=\textwidth]{figure/blood_ex_ggnet2-1.pdf}
\end{subfigure}
\begin{subfigure}[t]{.32\textwidth}
\caption{geom\_net}
\includegraphics[width=\textwidth]{figure/blood_ex_ggnet-1.pdf}
\end{subfigure}
\begin{subfigure}[t]{.32\textwidth}
\caption{ggnetwork}
\includegraphics[width=\textwidth]{figure/blood_ex_ggnetwork-1.pdf}
\end{subfigure}
\caption{\label{fig.cap:blood_ex} Network of blood donation possibilities in humans by ABO and RhD blood types.}
\end{figure}
\afterpage{\clearpage}
